\documentclass[runningheads,a4paper]{llncs}

\input{../preamble}

\input{../header}
\papertrue 
\techreportfalse
\newcommand{\mysectionref}{\iftechreport Chapter\else Section\fi}
\newcommand{\mysectionSref}{\iftechreport Chapters\else Sections\fi}
\newcommand{\mysectionrefnormal}{\iftechreport chapter\else section\fi}
\newcommand{\mychapter}[1]{\iftechreport \chapter{#1}\else \section{#1}\fi}
\newcommand{\mysection}[1]{\iftechreport \section{#1}\else \subsection{#1}\fi}
\newcommand{\mysubsection}[1]{\iftechreport \subsection{#1}\else \subsubsection{#1}\fi}

\bibliographystyle{abbrv} 
\usepackage{cite}

\begin{document}

\title{Many Suspensions, Many Problems:
A Review \\
 of Self-Suspending Tasks in Real-Time Systems}
\titlerunning{Review of Self-Suspending Tasks in Real-Time Systems}
\author{Jian-Jia Chen\inst{1}, Geoffrey Nelissen\inst{2}, Wen-Hung
  Huang\inst{1}, Maolin Yang\inst{3}, Bj\"orn Brandenburg\inst{4},
  Konstantinos Bletsas\inst{2}, Cong Liu\inst{5}, Pascal
  Richard\inst{6}, \\Fr\'ed\'eric Ridouard\inst{6}, Neil
  Audsley\inst{7},  
Raj Rajkumar\inst{8},
 Dionisio de Niz\inst{9}, and
Georg von der Br\"uggen\inst{1}}
\institute{TU Dortmund University, Germany\\
\and
%2
CISTER/INESC-TEC, ISEP, Polytechnic Institute of Porto, Portugal \\
\and
%3
University of Electron. Science and Technology of China, China\\
\and
%4
Max Planck Institute for Software Systems (MPI-SWS), Germany\\
\and
%5
University of Texas at Dallas, USA\\
\and
%6
LIAS/University of Poitiers, France\\
\and
%7
University of York, UK\\
\and
%8
Carnegie Mellon University, USA
\and
Software Engineering Institute (SEI), USA\\
%9
}

\authorrunning{Chen et al.}
\maketitle  
 
\input{../abstract} 
\input{../Introduction_}
\input{../Model_bis}
%\input{../Model}
\input{../Existing-Solutions}
\input{../Misconception}
\input{../Synchronization}
\input{../soft-realtime}
\input{../Hardness}
\input{../Summary}

\section*{Acknowledgments}

This paper has been supported by DFG, as part of the Collaborative
Research Center SFB876 (http://sfb876.tu-dortmund.de/).

This work was partially supported by National Funds through FCT/MEC (Portuguese Foundation for Science and Technology) and co-financed by ERDF (European Regional Development Fund) under the PT2020 Partnership, within project UID/CEC/04234/2013 (CISTER Research Centre); also by ARTEMIS/0003/2012 - JU grant nr. 333053 (CONCERTO) and ARTEMIS/0001/2013 - JU grant nr. 621429 (EMC2)

This material is based upon work funded and supported by NSF grants OISE 1427824 and CNS 1527727, and the Department of Defense under Contract No. FA8721-05-C-0003 with Carnegie Mellon University for the operation of the Software Engineering Institute, a federally funded research and development center.

\medskip
\noindent [Distribution Statement A] This material has been approved for public release and unlimited distribution. Please see Copyright notice for non-US Government use and distribution.

\medskip
\noindent Carnegie Mellon$^\copyright$ is registered in the U.S. Patent and Trademark Office by Carnegie Mellon University.

\medskip
\noindent DM-0003197

\bibliography{../bibliography/biblio}{}

\end{document}
